\LARGE \bf Predicting Initialization Effectiveness for Trajectory Optimization

### 2014-02-14 paper.tex
revised

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Trajectory optimization is a method for solving motion planning problems by formulating them as non-convex constrained optimization problems. The optimization process, however, can get stuck in local optima that are in collision. As a consequence, these methods typically require multiple initializations; this poses the problem of deciding which initializations to use, given a limited computational budget. In this paper we study the problem of predicting the quality of initial trajectories. Concretely, we propose a machine learning approach to predict whether a collision-free solution will be found from a given initialization. We present a set of trajectory features that encode the obstacle distribution locally around a robot. These features are designed such that they are transferable across different tasks. Experimental evaluation on various planning benchmarks demonstrates the accuracy of our approach when predicting the performance of state-of-the-art trajectory optimization---generalizing over tasks and environments.
\end{abstract}

@@

We propose a learning-based approach to predict the quality of a trajectory as an initialization to the optimization-based motion planning. Here, a trajectory's quality is measured by whether the optimization problem converges to a collision-free trajectory when using the given trajectory as the initial guess. In other words, the quality is defined in a qualitative manner, i.e. `good' or `bad'.
For the learning algorithm, we first design a set of trajectory features, which are relevant to a trajectory's effectiveness as an initialization. These features are both task-independent and scene-independent, and hence they are transferable among different planning tasks. We use four types of features: 1) spatial signed distance (SSD) vectors for each robot link; 2) difference between SSD vectors for the same robot link in adjacent trajectory waypoints and for adjacent robot links in the same waypoint; 3) spherical harmonics transform of SSD vectors; and 4) features measuring the non-convexity of the constrained optimization problem in the trajectory neighborhood. These features encode the obstacle distribution locally around a robot trajectory and thus are essential for predicting a trajectory's behavior under a trajectory optimization algorithm. We calcuate these features for a set of randomly selected trajectories for different tasks in various benchmarks. We then run optimization-based motion planning using these trajectories as initial guesses and obtain a labeled dataset. In the dataset, each trajectory is augmented with a `good' or `bad' label according to whether the optimization converges to a collision-free trajectory. With this dataset as the training set, we learn a predict function for trajectory effectiveness. We evaluate the accuracy of the learned predict function by counting the ratio of correct predictions for a set of new trajectories on various benchmarks.

@@

For general non-convex optimization problems, the dependence on initialization is also a well-known challenge. A simple and popular solution is using multiple random initializations to help the algorithms escape from bad local optima. 
One recent work closely related with our work is~\cite{Cassioli:2012:MLG}, which uses machine learning to build the relationship between the starting point of an optimization algorithm and the objective value of the final outcome. The main difference is that our approach is for trajectory optimization, where generating a feasible trajectory is more important than decreasing the absolute cost of the objective function. Moreover, trajectory optimization can exploit workspace heuristics for selecting good initial guesses, and such heuristics are not available in general optimization problems. Another closely related line of work is the trajectory prediction~\cite{Jetchev:2013:FMP}, where learning algorithms are used to predict a good path in a new environment from a database of demonstrated trajectories. The predicted result may not be feasible and hence requires another trajectory optimization step to resolve all violations.

Prior work has considered designing trajectory features for various applications. For example, Bentivegna et al.~\cite{Bentivegna:2006:LST} and Stolle et al.~\cite{Stolle:2007:TPT} represented a library of trajectories by feature vectors for control policy transferring. 
Berenson et al.~\cite{Berenson:2012:RPP} designed several criteria and features to evaluate whether a trajectory from a path library is likely to be reused in a new planning scene after suitable repair operations. Trajectory features are also used to predict class labels of moving objects based on their trajectories~\cite{Lee:2008:TTC}. Moreover, features have been designed to compress trajectory libraries~\cite{Arikan:2006:CMC}. However, all these features only capture properties of a trajectory itself. In contrast, the features proposed in our paper encode the interaction between the trajectory and the surrounding environment, which is critical for predicting how fast a trajectory can be pushed away from the obstacles by a trajectory optimization algorithm.

@@

Trajectory optimization is a challenging non-convex problem, and many approaches have been presented to solve it effectively. Among them are sequential convex optimization~\cite{Schulman:2013:FLO}, covariant Hamiltonian optimization~\cite{Ratliff:2009:CGO}, and stochastic optimization~\cite{Kalakrishnan:2011:STOMP}. However, given an initial trajectory that contains collisions, these methods are not guaranteed to find a collision-free solution due to the non-convex constraints in the optimization. Figure~\ref{fig:badgradient} shows some scenarios illustrating how trajectory optimization tends to get stuck in local optima that are not collision-free.

It is an on-going line of research to reshape the non-convex optimization formulations to improve their performance. In this paper we consider the complimentary problem of: Given a trajectory optimization approach, how to predict which initializations will result in collision-free solutions and which ones will not. Such predictive capabilities would enable selecting a good initial trajectory for motion planning so that trajectory optimization is likely to converge to a collision-free local optimum. Such information can also be used to improve the performance of trajectory optimization being run in parallel for multiple initializations: For each solution sequence starting with a different initial guess, we can stop bad solution sequences early enough and focus computational budgets on solution sequences that potentially can converge to good outcomes.

\subsection{Outline of Proposed Approach}
The problem to predict whether a given trajectory is a good or bad initial guess for a trajectory optimization algorithm can be formalized as a two-class classification problem. In fact, a trajectory is good if and only if it locates in the basin of attraction of a local optima that is collision-free, as shown in Figure~\ref{fig:overview}. In this way, all the trajectories in the trajectory space can be categorized into two classes, according to whether they are attracted to a collision-free or an in-collision local optimum.

\begin{figure}[!h]
\centering
\includegraphics[width=\linewidth]{figure/overview.pdf}
\caption{Trajectory quality prediction as a two-class classification problem. The curve is the trajectory cost manifold: the blue part corresponds to the collision-free trajectories while the black parts correspond to in-collision ones. The trajectory space can be classified into two classes (red and green) according to the basin of attraction that each trajectory locates in. }
\label{fig:overview}
\end{figure}

@@

Given a set of $N$ trajectories, we build a training set $\{\fsym_i, y_i\}_{i=1}^N$. Based on the data set, we learn a classifier $C(\fsym) = \{0, 1\}$, which is the predicator for a trajectory's effectiveness. Given a new trajectory as input, $C(\cdot)$ will provide its qualitative evaluation (good or bad) as output. The computed predicator $C(\cdot)$ may change among different planning methods, because different optimization approaches may label trajectories in the training set differently. However, our learning method is general and works for all trajectory optimization based planning methods.

\section{Predicting Initialization Effectiveness for Trajectory Optimization}
\label{sec:predict}

In this section, we design trajectory features to distinguish good and bad initializations and use the extracted features for predicting the effectiveness of a new trajectory.

Before extracting features from a trajectory, we first perform re-sampling on it so that waypoints are uniformly distributed between two end-points.
%We perform this pre-processing based on what was pointed out by Ratliff et al.~\cite{Ratliff:2009:CGO}: the objective function should be invariant to the re-parameterization of the trajectory so that the gradient will pull a trajectory out of obstacles instead of encouraging it to quickly jump through obstacles. Some recent work such as~\cite{Schulman:2013:FLO} handles the `jumping through problem' in a completely different way, but in practice, it is observed that a re-sampling pre-processing can still result in better convergence. 
The re-sampling process excludes the difference of trajectories in parameterization and hence makes it easier to compare them. After re-sampling, each trajectory has $T$ waypoints. This also guarantees our feature extraction gives fixed-length vectors for different trajectories.

%\subsection{Background: Signed Distances}
%We design features to encode the obstacle distribution locally around a trajectory, as such information is closely related with a trajectory's effectiveness. Intuitively speaking, if a trajectory is deep inside the obstacles, it is less likely to be pushed out of them; whereas if it is mostly outside the obstacles, it should converge to a collision-free solution. To quantify the obstacle distribution surrounding a trajectory, we use signed distances and normals, which are used in previous trajectory optimization algorithms such as~\cite{Schulman:2013:FLO}. Formally, the signed distance is defined between two sets $\mathcal A$ and $\mathcal B$ in workspace as follows:
%\begin{align}
%\text{sd}(\mathcal A, \mathcal B) = \text{dist}(\mathcal A, \mathcal B) - \text{penetration}(\mathcal A, \mathcal B).
%\end{align}
%Here $\text{dist}(\mathcal A, \mathcal B)$ is the distance between $\mathcal A$ and $\mathcal B$, i.e., the length of the smallest relative translation $\mathbf T$ that puts these two non-overlapped sets in contact:
%\begin{align}
%\text{dist}(\mathcal A, \mathcal B) = \inf\{\|\mathbf T\| \mid (\mathbf T + \mathcal A) \cap \mathcal B \neq \emptyset\}.
%\end{align}
%And $\text{penetration}(\mathcal A, \mathcal B)$ is the (translational) penetration depth between $\mathcal A$ and $\mathcal B$, i.e., the minimum translation $\mathbf T$ that makes two overlapped sets out of contact:
%\begin{align}
%\text{penetration}(\mathcal A, \mathcal B) = \inf\{\|\mathbf T\| \mid (\mathbf T + \mathcal A) \cap \mathcal B = \emptyset\}.
%\end{align}
%The signed distance computation can be efficiently computed for convex objects using GJK algorithm~\cite{Gilbert:1988:GJK} and EPA algorithm~\cite{Bergen:2001:EPA}. Signed distance also provides a normal vector $\mathbf n = \frac{\mathbf T}{\|\mathbf T\|}$, i.e., the direction that two objects can be most efficiently pushed away from each other. Such a direction is important for a trajectory's effectiveness: As we showed in Figure~\ref{fig:badgradient}, a bad initialization tends to have adjacent waypoints with very different pushing away directions.

\subsection{Background: Signed Distances}
We design features to encode the obstacle distribution locally around a trajectory, as such information is closely related with a trajectory's effectiveness. Intuitively speaking, if a trajectory is deep inside the obstacles, it is less likely to be pushed out of them; whereas if it is mostly outside the obstacles, it should converge to a collision-free solution. To quantify the obstacle distribution surrounding a trajectory, we use signed distances and normals, which are used in previous trajectory optimization algorithms such as~\cite{Schulman:2013:FLO}. Informally, signed distance between two objects is the length of the smallest translation that puts them in contact; signed normal is the direction of such translation. The signed distance computation can be efficiently computed for convex objects using GJK algorithm~\cite{Gilbert:1988:GJK} and EPA algorithm~\cite{Bergen:2001:EPA}.

\subsection{Feature I: Spatial Signed Distance Vectors}
Based on the signed distances formulation, a natural choice for designing trajectory feature is to collect signed distances and normals between all pairs of robot links and environment obstacles, for each waypoint in the trajectory. More formally, given a robot with $n$ links $\mathcal A_i$ in an environment with $m$ obstacles $\mathcal B_j$, the feature vector would be $\fsd = \{\text{sd}_{i,j,t}, \mathbf n_{i,j,t}\}_{1\leq i \leq n, 1\leq j \leq m, 1\leq t \leq T}$, where $\text{sd}_{i,j,t} = \text{sd}(\mathcal A_i(t), \mathcal B_j)$ is the signed distance value between $\mathcal A_i$ and $\mathcal B_j$ at waypoint $\btheta_t$, and $\mathbf n_{i,j,t}$ is the corresponding normal. This feature vector is of dimension $4\times m \times n \times T$ and provides a complete description for the obstacle distribution around the trajectory. However, the dimension of this feature depends on $m$, the number of obstacles in the environment, which makes this feature not transferable among different environments. Moreover, in many real-world systems, there may be a large number of obstacles (e.g., the obstacles from sensor data are usually represented as thousands of boxes~\cite{Sucan:2010:CPT}). This would result in a long feature vector and brings challenges for both data storage and the following learning procedure. Thus we choose to summarize this feature in a more compact manner but keep its rich information as much as possible.

@@

Our solution is to re-organize the signed distances and normals into a new form called the Spatial Signed Distance (SSD) Vectors. We illustrate the basic idea of SSD via a 2D example in Figure~\ref{fig:trajectoryfeaturePD}(a). We first partition the space around a robot link using a dictionary of direction vectors $D = \{\mathbf b_1, ..., \mathbf b_M\}$. These vectors are specially chosen so that they can divide the circle into bins with the same size. For 3D case, we generate the direction dictionary using the approach introduced in~\cite{Yershova:2010:GUI}, which computes a uniform deterministic sequence of samples over $S^2$ (as shown in Figure~\ref{fig:trajectoryfeaturePD}(b)) using a special multi-resolution grid structure. These dictionary directions partition sphere surface into bins with the same size.

Given the signed distances and normals between the robot link and all the obstacles in the environment, we accumulate the signed distances along these discretized prototype directions in dictionary $D$ and the result is a $M$-dim vector $\fssd$. The $k$-th element of $\fssd$ is denoted as $\fssd[k]$, which measures how hard it is to push the link away from the obstacle along direction $\mathbf b_k$. If $\fssd[k]$ is positive, then the link still has some distance from the obstacle and is safe; if $\fssd[k]$ is negative, then the link will go deeper inside the obstacle when being pushed in this direction and hence should be pushed in other directions.

@@

\subsection{Feature II: Difference between Spatial Signed Distance Vectors}
Spatial signed distance feature provides a complete description about the obstacle distribution around the trajectory. However, as we mentioned in Section~\ref{sec:probstatement}, whether the signed distance normal is consistent between adjacent links or adjacent waypoints in an initial trajectory is also important for the convergence of a trajectory optimization algorithm. Thus, we design features to measure the consistency between two spatial signed distance feature vectors. Given two vectors $\fssd^a$ and $\fssd^b$, we evaluate their consistency using their dot-product $\fssd^a \cdot \fssd^b$. This dot-product has a larger value when $\fssd^a[k]$ and $\fssd^b[k]$ are of the same sign for all $k$. And it has a smaller value when some $\fssd^a[k]$ and $\fssd^b[k]$ are of the opposite sign, i.e., the signed distance normals are not consistent in direction $\mathbf b_k$. We compute dot-product between SSD vectors of the same link in adjacent waypoints and the adjacent links in the same waypoint. We collect all these results in a feature vector denoted as $\fcon$, i.e., the consistency feature.

\subsection{Feature III: Spherical Harmonics Spatial Feature}
Spatial signed distance feature is not rotation-invariant and this may have some issues.
For example, when both the environment and the trajectory are rotated by a given angle, the trajectory feature will be completely different; and thus it would be difficult to transfer the learning result between different benchmarks. Another example is when the environment is symmetric (such as Figure~\ref{fig:benchmarks1}(c)), the lack of rotation-invariance in the feature may decrease the prediction accuracy of the learning algorithm.

As shown in Figure~\ref{fig:trajectoryfeaturePD}(a), the spatial signed distance feature in fact is a function defined on a sphere around a robot link. To represent the feature in a rotation-invariant manner, we utilize the mathematical concept of spherical harmonics. The same idea has been used in various applications, including computer vision~\cite{Frome:2004:ROR} and graphics~\cite{Kazhdan:2003:RIS}.

The theory of spherical harmonics says that any spherical function $s(\theta, \phi)$ can be expressed as a sum of complex spherical harmonic basis functions $Y_l^m$: 
\begin{equation}
s(\theta, \phi) = \sum_{l=1}^{\infty} \sum_{m=-l}^{l} a_l^m Y_l^m(\theta, \phi).
\end{equation}
According to the property of spherical harmonics transform, the amplitudes of the harmonic coefficients $\|a_l^m\|$ are invariant to any rotation in the azimuthal direction. Moreover, the amplitude of $s_l$, the $l$-th frequency component of the function $s$, is rotation-invariant under any rotations, where $s_l = \sum_{m=-l}^{l}a_l^m Y_l^m(\theta, \phi)$. The magnitude of $s_l$ can also be computed based on $a_l^m$: $\|s_l\|^2 = \sum_{m=-l}^l \|a_l^m\|^2$. We use $\|s_l\|$ as features in 3D benchmarks. In 2D case, we use $\|a_l^m\|$ as features since only azimuthal rotation is allowed in 2D and $\|a_l^m\|$ can provide richer information. Similar to previous work~\cite{Frome:2004:ROR,Kazhdan:2003:RIS}, we choose a bandwidth $b$ and store only $b$ lowest-frequency components in our spherical harmonics feature $\fsht$. More formally, the feature is defined as $\fsht = \{\|s_l\|\}$ in 3D and $\fsht = \{\|a_l^m\|\}$ in 2D, where $m = 0,...,l$ and $l = 0,...,b$.

@@

We first compute Hessian matrices of the objective and constraint functions at the trajectory. Next, to measure the local non-convexity of the optimization problem, we compute following features based on the eigenvalues of the Hessian matrices: 1) the minimum eigenvalue; 2) the sum of negative eigenvalues; 3) the maximum eigenvalue; 4) the sum of all eigenvalues. We compute these four statistics for each objective and constraint function, and generate a total of $4 \times (1 + n_{\text{ieq}} + n_{\text{eq}})$ convexity features, which are collected in a feature vector denoted as $\fconv$.

@@

Given a set of trajectories with their features and evaluation results, we can use any two-class classification algorithm to learn a classifier for the trajectory. As the dimension of our trajectory feature is high (larger than $200$), we choose the linear support vector machine~\cite{Fan:2008:LLL} as the classifier due to its simplicity and efficiency.

@@

Given one planning task, we first generate a trajectory which is a simple linear interpolation between the initial and goal configurations. Next, we perform random perturbation of the trajectory waypoints and generate more random trajectories. For each of these trajectories, we run the trajectory optimization algorithm using it as the initial guess. When the optimization stops, we will obtain a sequence of trajectories. These intermediate trajectories can also be used as initializer and they would all have the same effectiveness label as the actual initial guess. As a result, if the optimization converges and the final outcome is feasible, we evaluate the given initial trajectory and all intermediate trajectories as `good' initializations. Otherwise we evaluate all of them as `bad'. Finally, we extract features for all these labeled trajectories and add them into the dataset. In this way, we generate more than $10,000$ trajectories for each planning benchmark. For each benchmark, we use half of the data for training a classifier and the rest for testing experiments. When splitting training and test sets, we make sure that trajectories corresponding to the same task are assigned into the same set, in order to guarantee that the test set is independent with the training set. In Figure~\ref{fig:features}, we visualize the features of two trajectories from different benchmarks. Except $\fconv$, all these features can be computed in less than 10 milleseconds per trajectory. $\fconv$ is more expensive (several seconds) since it is time consuming to compute a trajectory's Hessian matrix numerically.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Trajectory optimization is a method for solving motion planning problems by formulating them as non-convex constrained optimization problems. The optimization process, however, can get stuck in local optima that are in collision. As a consequence, these methods typically require multiple initializations.  This poses the problem of deciding which initializations to use when given a limited computational budget. In this paper we propose a machine learning approach to predict whether a collision-free solution will be found from a given initialization. We present a set of trajectory features that encode the obstacle distribution locally around a robot. These features are designed for generalization across different tasks.  Our experiments on various planning benchmarks demonstrate the performance of our approach.
\end{abstract}

@@

We propose a learning-based approach to predict the quality of a trajectory as an initialization to the optimization-based motion planning. Here, a trajectory's quality is measured by whether the optimization problem converges to a collision-free trajectory when using the given trajectory as the initial guess. In other words, the quality is defined in a qualitative manner, i.e. `good' or `bad'.
For the learning algorithm, we first design a set of trajectory features, which are relevant to a trajectory's effectiveness as an initialization. These features are both task-independent and scene-independent, and hence they are transferable among different planning tasks. We use four types of features: (i) spatial signed distance (SSD) vectors for each robot link; (ii) difference between SSD vectors for the same robot link in adjacent trajectory waypoints and for adjacent robot links in the same waypoint; (iii) spherical harmonics transform of SSD vectors; and (iv) features measuring the non-convexity of the constrained optimization problem in the trajectory neighborhood. These features try to capture the obstacle distribution locally around a robot trajectory and thus can be expected to be useful for predicting a trajectory's behavior under a trajectory optimization algorithm. We calcuate these features for a set of randomly selected trajectories for different tasks in various benchmarks. We then run optimization-based motion planning using these trajectories as initial guesses and obtain a labeled dataset. In the dataset, each trajectory is augmented with a `good' or `bad' label according to whether the optimization converges to a collision-free trajectory. With this dataset as the training set, we learn a prediction function for trajectory effectiveness. We evaluate the accuracy of the learned prediction function by counting the ratio of correct predictions for a set of new trajectories on various benchmarks.

@@

For general non-convex optimization problems, the dependence on initialization is also a well-known challenge. A simple and popular solution is using multiple random initializations to help the algorithms escape from bad local optima. 
In recent work~\cite{Cassioli:2012:MLG} machine learning was used to build the relationship between the starting point of an optimization algorithm and the objective value of the final outcome. The main difference is that our approach is for trajectory optimization, where generating a feasible trajectory is more important than decreasing the absolute cost of the objective function. Moreover, trajectory optimization can exploit workspace heuristics for selecting good initial guesses, and such heuristics are not available in general optimization problems. Another closely related line of work is about trajectory prediction~\cite{Jetchev:2013:FMP}, where learning algorithms are used to predict a good path in a new environment from a database of demonstrated trajectories. The predicted result may not be feasible and hence requires a second step to resolve all violations.

Prior work has considered designing trajectory features for various applications. For example, Bentivegna et al.~\cite{Bentivegna:2006:LST} and Stolle et al.~\cite{Stolle:2007:TPT} represented a library of trajectories by feature vectors for control policy transfer. 
Berenson et al.~\cite{Berenson:2012:RPP} designed several criteria and features to evaluate whether a trajectory from a path library is likely to be reused in a new planning scene after suitable repair operations. Trajectory features are also used to predict class labels of moving objects based on their trajectories~\cite{Lee:2008:TTC}. Moreover, features have been designed to compress trajectory libraries~\cite{Arikan:2006:CMC}. However, all these features only capture properties of a trajectory itself. In contrast, the features proposed in our paper encode the interaction between the trajectory and the surrounding environment, which is critical for predicting how fast a trajectory can be pushed away from the obstacles by a trajectory optimization algorithm.

@@

Trajectory optimization is a challenging non-convex problem, and many approaches have been presented to solve it effectively. Among them are sequential convex optimization~\cite{Schulman:2013:FLO}, covariant Hamiltonian optimization~\cite{Ratliff:2009:CGO}, and stochastic optimization~\cite{Kalakrishnan:2011:STOMP}. However, given an initial trajectory that contains collisions, none of these methods are guaranteed to find a collision-free solution due to the non-convex constraints in the optimization. Figure~\ref{fig:badgradient} shows some scenarios illustrating how trajectory optimization tends to get stuck in local optima that are not collision-free.

It is an on-going line of research to reshape the non-convex optimization formulations to improve their performance. In this paper we consider the following complimentary problem: Given a trajectory optimization approach, how to predict whether an initialization will result in a collision-free solution or not. 
Such predictive capabilities would enable selecting a good initial trajectory for motion planning so that trajectory optimization is likely to converge to a collision-free local optimum. Such information can also be used to improve the performance of trajectory optimization being run in parallel for multiple initializations: For each solution sequence starting with a different initial guess, we can stop bad solution sequences early enough and focus computational budgets on solution sequences that potentially can converge to good outcomes.

\subsection{Outline of Proposed Approach}
The problem to predict whether a given trajectory is a good or bad initial guess for a trajectory optimization algorithm can be formalized as a two-class classification problem. In fact, a trajectory is good if and only if it locates in the basin of attraction of a local optimum that is collision-free.

@@

Given a set of $N$ trajectories, we build a training set $\{\fsym_i, y_i\}_{i=1}^N$. Based on the data set, we learn a classifier $C(\fsym) $, which is the predictor for a trajectory's effectiveness. Given a new trajectory as input, $C(\cdot)$ will provide its qualitative evaluation (good or bad) as output. The learned predictor $C(\cdot)$ can depend on the planning methods, because different optimization approaches may differ in for which initial trajectories they end up finding feasible solutions.  However, our learning method is general and does not depend on the type of trajectory optimization method.

\section{Predicting Initialization Effectiveness for Trajectory Optimization}
\label{sec:predict}
In this section, we design trajectory features to distinguish good and bad initializations and use the extracted features for predicting the effectiveness of a new trajectory.

Before extracting features from a trajectory, we first perform re-sampling on it so that waypoints are uniformly distributed between two end-points.
We perform this pre-processing based on what was pointed out by Ratliff et al.~\cite{Ratliff:2009:CGO}: the objective function should be invariant to the re-parameterization of the trajectory so that the gradient will pull a trajectory out of obstacles instead of encouraging it to quickly jump through obstacles. Some recent work such as~\cite{Schulman:2013:FLO} handles the `jumping through problem' in a completely different way, but in practice, it is observed that a re-sampling pre-processing can still result in better convergence. 
The re-sampling process removes some the parameterization-dependent differences between trajectories and hence makes it easier to compare them. After re-sampling, each trajectory has $T$ waypoints. This also results in fixed-length feature vectors.

\subsection{Background: Signed Distances}
We design features to encode the obstacle distribution locally around a trajectory, as such information is closely related with a trajectory's effectiveness. Intuitively speaking, if a trajectory is deep inside the obstacles, it is less likely to be pushed out of them; whereas if it is mostly outside the obstacles, it should converge to a collision-free solution. To quantify the obstacle distribution surrounding a trajectory, we use signed distances and normals, which are used in previous trajectory optimization algorithms such as~\cite{Schulman:2013:FLO}. Informally, the signed distance between two objects is the length of the smallest translation that puts them in contact; the signed normal is the direction of such translation. The signed distance can be efficiently computed for convex objects using GJK algorithm~\cite{Gilbert:1988:GJK} and EPA algorithm~\cite{Bergen:2001:EPA}.   For use with non-convex objects, these objects can be represented as a set of convex objects together making up the original non-convex object.

\subsection{Feature I: Spatial Signed Distance Vectors}
Based on the signed distances formulation, a natural choice for designing trajectory feature vector is to collect signed distances and normals between all pairs of robot links and environment obstacles, for each waypoint in the trajectory. More formally, given a robot with $n$ links $\mathcal A_i$ in an environment with $m$ obstacles $\mathcal B_j$, the feature vector would be $\fsd = \{\text{sd}_{i,j,t}, \mathbf n_{i,j,t}\}_{1\leq i \leq n, 1\leq j \leq m, 1\leq t \leq T}$, where $\text{sd}_{i,j,t} = \text{sd}(\mathcal A_i(t), \mathcal B_j)$ is the signed distance value between $\mathcal A_i$ and $\mathcal B_j$ at waypoint $\btheta_t$, and $\mathbf n_{i,j,t}$ is the corresponding normal. This feature vector is of dimension $4\times m \times n \times T$ and provides a complete description for the obstacle distribution around the trajectory. However, the dimension of this feature depends on $m$, the number of obstacles in the environment, which makes this feature not transferable among different environments. Moreover, in many real-world systems, there may be a large number of obstacles (e.g., the obstacles from sensor data are usually represented as thousands of boxes~\cite{Sucan:2010:CPT}). This would result in a long feature vector and brings challenges for both data storage and the following learning procedure. Thus we propose to summarize this information in a more compact manner that lends itself better to generalization.

@@

Our solution is to re-organize the signed distances and normals into a new form called the Spatial Signed Distance (SSD) vectors. We illustrate the basic idea of SSD via a 2D example in Figure~\ref{fig:trajectoryfeaturePD}(a). We first partition the space around a robot link using a dictionary of direction vectors $D = \{\mathbf b_1, ..., \mathbf b_M\}$.  For the 3D case, we generate the direction dictionary using the approach introduced in~\cite{Yershova:2010:GUI}, which computes a uniform deterministic sequence of samples over $S^2$ (as shown in Figure~\ref{fig:trajectoryfeaturePD}(b)) using a specific multi-resolution grid structure.  This partitions the sphere surface into bins of the same size.

Given the signed distances and normals between the robot link and all the obstacles in the environment, we accumulate the signed distances along these discretized prototype directions in dictionary $D$ and the result is a $M$-dimensional vector $\fssd$. The $k$-th element of $\fssd$ is denoted as $\fssd[k]$, which summarizes signed distances along direction $\mathbf b_k$. If $\fssd[k]$ is positive, then the link still has some distance from the obstacle and is safe; if $\fssd[k]$ is negative, then the link will go deeper inside the obstacle when being pushed in this direction and hence should be pushed in other directions.

@@

\subsection{Feature II: Difference between Spatial Signed Distance Vectors}
The spatial signed distance vector provides a description of the obstacle distribution around the trajectory on a per link basis. However, as we mentioned in Section~\ref{sec:probstatement}, whether the signed distance normal is consistent between adjacent links or adjacent waypoints in an initial trajectory is also important for the convergence of a trajectory optimization algorithm. Thus, we include features to measure the consistency between two spatial signed distance feature vectors. Given two vectors $\fssd^a$ and $\fssd^b$, we evaluate their consistency using their dot-product $\fssd^a \cdot \fssd^b$. This dot-product has a larger value when $\fssd^a[k]$ and $\fssd^b[k]$ are of the same sign for all $k$. And it has a smaller value when some $\fssd^a[k]$ and $\fssd^b[k]$ are of the opposite sign, i.e., the signed distance normals are not consistent in direction $\mathbf b_k$. We compute dot-product between SSD vectors of the same link in adjacent waypoints and between adjacent links in the same waypoint. We collect all these results relating to consistency in the signed distances in a feature vector denoted by $\fcon$.

\subsection{Feature III: Spherical Harmonics Spatial Feature}
The spatial signed distance feature is not rotation-invariant.
As a result when, for example, both the environment and the trajectory are rotated by a given angle, the feature vector will be completely different.  This could affect its generalization capability. Another example is when the environment is symmetric (such as Figure~\ref{fig:benchmarks1}(c)).  Overall, the lack of rotation-invariance in the feature may decrease the prediction accuracy of the learning algorithm.

As shown in Figure~\ref{fig:trajectoryfeaturePD}(a), the spatial signed distance feature is a function defined on a sphere around a robot link. To represent the feature in a rotation-invariant manner, we utilize spherical harmonics. The same idea has been used in various applications, including computer vision~\cite{Frome:2004:ROR} and graphics~\cite{Kazhdan:2003:RIS}.

The theory of spherical harmonics says that any spherical function $s(\theta, \phi)$ can be expressed as a sum of complex spherical harmonic basis functions $Y_l^m$: 
\begin{equation}
s(\theta, \phi) = \sum_{l=1}^{\infty} \sum_{m=-l}^{l} a_l^m Y_l^m(\theta, \phi).
\end{equation}
The amplitudes of the harmonic coefficients $\|a_l^m\|$ are invariant to any rotation in the azimuthal direction. Moreover, the amplitude of $s_l$, the $l$-th frequency component of the function $s$, is rotation-invariant under any rotations, where $s_l = \sum_{m=-l}^{l}a_l^m Y_l^m(\theta, \phi)$. The magnitude of $s_l$ can also be computed based on $a_l^m$: $\|s_l\|^2 = \sum_{m=-l}^l \|a_l^m\|^2$. We use $\|s_l\|$ as features in 3D benchmarks. In 2D case, we use $\|a_l^m\|$ as features since only azimuthal rotation is allowed in 2D and $\|a_l^m\|$ can provide richer information. Similar to previous work~\cite{Frome:2004:ROR,Kazhdan:2003:RIS}, we choose a bandwidth $b$ and store only $b$ lowest-frequency components in our spherical harmonics feature $\fsht$. More formally, the feature is defined as $\fsht = \{\|s_l\|\}$ in 3D and $\fsht = \{\|a_l^m\|\}$ in 2D, where $m = 0,...,l$ and $l = 0,...,b$.

@@

We first compute Hessian matrices of the objective and constraint functions at the trajectory. Next, to measure the local non-convexity of the optimization problem, we compute the following features based on the eigenvalues of the Hessian matrices: (i) the minimum eigenvalue; (ii) the sum of negative eigenvalues; (iii) the maximum eigenvalue; (iv) the sum of all eigenvalues. We compute these four statistics for each objective and constraint function, and generate a total of $4 times T \times (1 + n_{\text{ieq}} + n_{\text{eq}})$ convexity features, which are collected in a feature vector denoted as $\fconv$.

@@

Given a set of trajectories with their features and evaluation results, we can use any two-class classification algorithm to learn a classifier for the trajectory. As the dimension of our trajectory feature is high (larger than $200$), we choose the linear support vector machine~\cite{Cortes:1995:SN} as the classifier due to its simplicity and efficiency.

@@

Given one planning task, we first generate a trajectory which is a simple linear interpolation between the initial and goal configurations. Next, we perform random perturbation of the trajectory waypoints and generate more random trajectories. For each of these trajectories, we run the trajectory optimization algorithm using it as the initial guess. When the optimization stops, we will obtain a sequence of trajectories. These intermediate trajectories can also be used as initializer and they would all have the same effectiveness label as the actual initial guess. As a result, if the optimization converges and the final outcome is feasible, we evaluate the given initial trajectory and all intermediate trajectories as `good' initializations. Otherwise we evaluate all of them as `bad'. Finally, we extract features for all these labeled trajectories and add them into the dataset. In this way, we generate more than $10,000$ trajectories for each planning benchmark. For each benchmark, we use half of the data for training a classifier and the rest for testing experiments. When splitting training and test sets, we make sure that trajectories corresponding to the same task are assigned into the same set, in order to guarantee that the test set is independent of the training set. In Figure~\ref{fig:features}, we visualize the features of two trajectories from different benchmarks. Except $\fconv$, all these features can be computed in less than 10 milleseconds per trajectory. $\fconv$ is more expensive (several seconds) since it is time consuming to compute a trajectory's Hessian matrix numerically.
